### 2017-01-31 inst/examples/joss.Rmd
Documents for JOSS Submission Prep.

---
title: "Rmd to HTML"
author: "Brandon LeBeau"
date: "January 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simple Markup
The `highlightHTML` packages allows for {#bggrey simple markup} to add formatting to text and tables. Using the `rmarkdown` package and the `table_id_inject` function allows users to easily add markup to Rmd documents and render directly to HTML.

```{r symm, echo = FALSE}
library(dplyr)
library(highlightHTML)

chick_tab <- chickwts %>%
  group_by(feed) %>%
  summarise(avg_weight = mean(weight),
            sd_weight = sd(weight))

table_id_inject(chick_tab, id = c('#bggrey', '#bgblack'), 
                conditions = c('> 270', '> 300'),
                variable = list('avg_weight', 'avg_weight'))
```