### 2017-11-29 inst/examples/joss.Rmd

---
title: "Rmd to HTML"
author: "Brandon LeBeau"
date: "January 13, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simple Markup
The `highlightHTML` packages allows for {#bggrey simple markup} to add formatting to text and tables. Using the `rmarkdown` package and the `table_id_inject` function allows users to easily add markup to Rmd documents and render directly to HTML.

```{r symm, echo = FALSE, results = 'asis', message = FALSE}
library(dplyr)
library(highlightHTML)

chickwts %>%
  group_by(feed) %>%
  summarise(avg_weight = mean(weight),
            sd_weight = sd(weight)) %>%
  mutate(feed = as.character(feed)) %>%
  table_id_inject(id = c('#bggrey', '#bgblack', '#bglightred', '#textblue'), 
                conditions = c('> 270', '> 300', '> 60', '== "horsebean"'),
                variable = list('avg_weight', 'avg_weight', 'sd_weight', 'feed'),
                num_digits = 3) %>%
  knitr::kable(format = 'markdown')
```